---
title: "Bird Joins"
author: "Scout Leonard"
date: "8/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)


bird_observations <- read_csv(here::here("data","knb-lter-cap", "52_pp52_birds_1.csv"))

sites <- read_csv(here::here("data", "knb-lter-cap", "52_pp52_sites_1.csv"))

surveys <- read_csv(here::here("data", "knb-lter-cap", "52_pp52_surveys_1.csv"))

taxalist <- read_csv(here::here("data", "knb-lter-cap", "52_pp52_taxalist_1.csv"))
```


##Task 1: Joins on birds

**Create a subset of `bird_observations` called `birds_subset` that only contains observation for birds with species id "BCHO" and "RWBL." and from sites with ID "LI-W" and "NU-C"

```{r}
birds_subset <- bird_observations %>% 
  filter(species_id %in% c("BHCO", "RWBL"),
         site_id %in% c("LI-W", "NU-C"))
```

### Left Join practice

```{r}
birds_left <- left_join(birds_subset, sites, by = "site_id") 

birds_left_1 <- left_join(birds_left, taxalist, by = "species_id")
```

### Full Join practice 

I expect a `full_join()` between `birds_subset` and `sites` to contain all of the observations in the first df, the second, and both, with NA values for variables missing from observations in one or the other. 

```{r}
birds_full <- full_join(birds_subset, sites, by = "site_id")
```

### Inner join practice

I expect an `inner_join()` between `birds_subset` and `taxalist` to contain all of the observation from both dfs. 

```{r}
birds_inner <- inner_join(birds_subset, taxalist, by = "species_id")
```

If instead of `inner_join()` I used `left_join()`, I would expect the data frame to include all of the elements from the birds subset df with additional data from the elements that match in the taxalist df. 
## Task 2: Data Wrangling and visualization with birds 

```{r}
bird_observations_1 <- bird_observations %>% 
  rename(bird_obs_notes = notes)

all_obs_birds_1 <- full_join(bird_observations_1, taxalist, by = "species_id")

all_obs_birds_2 <- full_join(all_obs_birds_1, sites, by = "site_id")

all_obs_birds_3 <- full_join(all_obs_birds_2, surveys, by = c("site_id", "survey_id"))

all_obs_birds <- all_obs_birds_3 %>% 
  select(survey_date, common_name, park_name, bird_count)
```

Use `lubridate::month()` to add a new column called `survey_month`, containing only the month number. Then convert the month to a factor (again with `mutate()`)

```{r}
all_obs_birds_month <- all_obs_birds %>% 
  mutate(survey_month = lubridate::month(survey_date)) %>% 
  mutate(survey_month = as.factor(survey_month)) %>% 
  dplyr::relocate(survey_month, .after = survey_date)

total_birds <- all_obs_birds_month %>% 
  group_by(park_name, survey_month) %>% 
  summarize(total_count = sum(bird_count, na.rm = TRUE)) %>% 
  filter(park_name %in% c("Lindo","Orme","Palomino","Sonrisa"))
```


